---
title: "Advanced Data Analytics"
author:
  - Loukas Gonzalez
  - Simon Alvarez Bliffeld
  - Simon Gonzalez
  - Yannik Biebert
date: "2023-11-28"
output:      
  rmarkdown::html_document:
    theme: cosmo
    highlight: tango
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r setup, include=FALSE}
# Set up: do not change!!!
knitr::opts_chunk$set(echo = TRUE)
```

<br>

------------------------------------------------------------------------

<br>

For the final project of Advanced Data Analytics, we decided to analyze three variables of two different countries, and with them, find and dissect valuable information, with which we can after create forecasts. The two countries that we decided on were Italy and the US, two countries that despite their distinct economic landscapes, share a profound impact on the global stage. The variables that we will be using will be the country's GDP, which is one of the most crucial indicators of the overall growth of the country, the CPI or Consumer Price Index, which goes along with the GDP but also gives us further information, and the Stock Market Index for each country. 

# Part 1: Prepare and inspect the data

-   Visual inspection of data and season, trend DONE, including desriptions


Steps for us:


-   Stationarity testing DONE

-   Inspect autocorrelation with acf and pacf DONE

-   Inspect volatility clustering (if so, maybe GARCH needed?)

-   if it has season use holt-winters, DO WE NEED TBATS or STLM? Because I dont think whe have two cicles in each other

-   Maybe new autocorrelation with acf and pacf for the estimated seasonality and trend (if we use the holt winter model to mode arima)

-   Look for ARIMA models (on data and on estimated ets()/msts()

-   Identify the seasonal model by analyzing the seasonal coefficients of the ACF and PACF

-   Identify the regular component by exploring the ACF and PACF of the residuals of the seasonal model.

-   Compare models (including ets models)- AIC or Accuracy Metrics

-   Choose best model and visualize, look at residuals, maybe GARCH (or should be aready dicide this before using ARIMA Model and use it directly?)

-   Use maybe Garch, and compare those models

-   Summary of findings

-   Forcast and plot final model

-   conclusion of forecast

------------------------------------------------------------------------

## 1.1 Importing and inspecting the data set

```{r Import data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# import packages
library(pacman)
p_load(
  readxl, tseries, tidyverse, dplyr, tidyr, forecast, fpp2, stats, quantmod, zoo, lubridate, ggplot2, cowplot
)

# import data set xlxs
data <- read_excel("Datasets.xlsx")

# removing not used columns Housing and Unemployment
data <- data[, -c(7, 9)]


# Display the data set
print(data)

```

As observed, the years that we will be using for this analysis are from 2000 to 2020, giving us a comprehensive analysis of each country's variables throughout the 21st century. 

## 1.2 Prepare the data


```{r transform data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Split the data into two dataframes based on the country
data_italy <- subset(data, Country %in% c("Italy"))
data_usa <- subset(data, Country %in% c("USA"))

# Function for data transformation
transform_data <- function(df) {
  # Extract columns from the dataframe
  gdp <- df$'GDP, billion currency units'
  cpi <- df$'Consumer Price Index (CPI)'
  stocks <- df$'Stock market index'

  # Drop missing values (becuase only every quarter)
  gdp <- na.omit(gdp)

  # Drop missing values (becuase last in the month)
  cpi <- na.omit(cpi)
  stocks <- na.omit(stocks)

  # Change to time series
  gdp <- ts(gdp, start = c(2000, 1), frequency = 4)
  cpi <- ts(cpi, start = c(2000, 1), frequency = 12)
  stocks <- ts(stocks, start = c(2000, 1), frequency = 12)
  
  # Drop everything after 2019
  gdp <- window(gdp, end = c(2019, 4))
  cpi <- window(cpi, end = c(2019, 12))
  stocks <- window(stocks, end = c(2019, 12))
  
  # Return the time series as separate variables
  return(list(gdp = gdp, cpi = cpi, stocks = stocks))
}

# Apply the transformation function
transformed_italy  <- transform_data(data_italy)
transformed_usa <- transform_data(data_usa)

```

After analyzing the data, we decided to omit the information from 2020 onwards. The Covid pandemic in 2020 significantly disrupted global economic patterns, causing unprecedented fluctuations in GDP, CPI, and Stock Market Indices. The widespread lockdowns, supply chain disruptions, and shifts in consumer behavior during the pandemic created outliers in the data that could skew our analysis and adversely affect the accuracy of our forecasts.

## 1.3 Visual inspection of the data, including Seasonality

```{r plot gdp in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Create a function for visualization.
plot_and_zoom <- function(data, variable, country,num_years) {
  
  # Extract the specified number of years from the time series
  end_date <- start(data) + c(num_years, 0)
  window_data <- window(data, start = start(data), end = end_date)
  
  # Create the main plot
  main_plot <- ggplot(data, aes(x = time(data), y = as.numeric(data))) +
    geom_line() +
    theme_minimal() +
    labs(title = paste0(variable, ' in ', country), x = "Year", y = variable) +
    theme(panel.grid.minor = element_blank(),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
  
  # Create a zoomed-in version with the specified number of years
  zoomed_in_plot <- ggplot(window_data, aes(x = time(window_data), y = as.numeric(window_data))) +
    geom_line() +
    labs(title = paste0('Zoomed in ',variable, ' in ', country), x = "Year", y = variable) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
  
  plot_grid(main_plot, zoomed_in_plot, ncol = 1)
  
}

# Plot the GDP for Italy
plot_and_zoom(transformed_italy$gdp, 'GDP', 'Italy', 4)

```

We can make some observations by looking at the graph of Italy's GDP from 2000 - 2020. The GDP shows a clear upward trend, with a sharper increase from 2000 to 2008, where it stabilizes, and again an increase from 2015-2020. The interruption in the trend corresponds to the 2008 financial crisis, which shook the economy and its growth. We can also note a seasonality throughout the time series. Italy's economy after 2001 began to slow down, even though the GDP is seen to be growing, the rate of growth had decreased. Factors such as "aging population, inflation, stagnation, and the impact of globalization on trade and demand" heavily affected the country's development and economic growth (Ardeni & Gallegati, 2023). Italy's development is slow mainly when compared to other countries European countries with similar capabilities and factors. We can observe another factor in the zoomed version of Italy's GDP. The plot shows a clear trend or pattern of seasonality, wherein in the first part of the year we observe a slight increase in the GDP, followed by a sharper increase in the second part of the year, and this is observed in all of the years included. This could be attributed to factors such as traditional vacation patterns, tourism peaks, or specific industry behaviors.

```{r plot cpi in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot the CPI for Italy
plot_and_zoom(transformed_italy$cpi, 'CPI', 'Italy', 4)


```

Italy’s CPI has seen a similar path to the GDP. From 2000 onwards, Italy's Consumer Price Index (CPI) has undergone fluctuations influenced by various economic factors. In the early 2000s, the CPI experienced constant growth, which continued until 2012, when it began to slow down and reduce the growth rate. This might be the consequence on the financial crisis, combined with internal economic and political issues. The zoomed plot also reveals seasonality, as we can observe dips in the CPI in the first part of each year for all of the observed periods. This could be attributed to high spending seasons of the year, as it usually is found in most CPI plots, such as holidays. 

```{r plot stocks in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot stocks for Italy
plot_and_zoom(transformed_italy$stocks, 'Stocks', 'Italy', 4)

```

The Stock Market Index for Italy is very different from the previous charts and datasets. As observed, the market had many fluctuations throughout the 2000 - 2020 period. We can observe that from 2000 to 2003, Italy’s stock market had a steep decline, followed by a rapid growth period that was again stopped by the financial crisis of 2008-2009. The index maintained high volatility for the following period, with short periods of growth or decline, but with an overall positive growth trend. Again, most of the behavior of the market is related to political decisions that the Italian government has made and external economic factors influencing investor confidence. Notably, the stock market showed increased volatility during times of political unpredictability or changes in governmental policy. Furthermore, the overall state of the world economy, particularly the Eurozone crisis, had a big impact on how the Italian stock market developed (Codogno & Galli, 2022). 

```{r plot gdp in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot GDP for USA
plot_and_zoom(transformed_usa$gdp, 'GDP', 'USA', 4)

```

In the United States GDP time series, we again observe a positive trend, with periods of faster increase. There is a specific moment where we see an interruption in the growth, which is in 2008, corresponding to the Financial Crisis. After this period, we can observe that the GDP grew at a faster rate. Aiming to revive the economy, major government interventions, fiscal stimulus packages, and monetary policies are responsible for these accelerations in GDP growth in the post-crisis decades. The United States has gained more and more power and importance globally, which again also influenced the constant growth of its GDP (Salvatore, 2020). No indications of seasonality or a specific pattern can be observed in the plot or the zoomed plot. Moreover, it can be observed that the GDP for USA is significantly higher than Italy’s, mostly as a result of the US doing more business abroad and having a bigger economy. The USA's economy is stronger than Italy's due in part to their larger corporations, advancements in technology, and international trading practices.  

```{r plot cpi in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot CPI for USA
plot_and_zoom(transformed_usa$cpi, 'CPI', 'USA', 4)

```

The USA’s CPI follows a very positive trend, with continuous growth throughout the period observed. We can observe certain moments of slowdown, specifically after the 2008 financial crisis, but even with this, the CPI is constantly growing, at a somewhat slow rate. No seasonality or specific pattern can be observed in the plot or the zoomed plot. This upward trend points to an economy that is typically growing and stable, with rare setbacks like the financial crisis eventually being resolved. These variations in the growth rate could also be a sign of the economy's resiliency and capacity to bounce back from setbacks, which would explain the CPI's general upward tendency.

```{r plot stocks in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot stocks for USA
plot_and_zoom(transformed_usa$stocks, 'Stocks', 'USA', 4)

```

The Stock Market index for the USA different from what we observed in Italy, shows a clear upward trend, with some periods of higher volatility and shifts. Again, as noted in all of the previous plots, we see a period of decline right in 2008, which corresponds to the financial crisis, where gains were lost in a very short period. The U.S. stock market proved strong after this trying time, rising and showing a steady growth trend, confirming the economy's capacity to come back from difficult periods. The U.S. financial system's inherent flexibility and the proactive actions of policymakers are frequently credited for this recovery.

## 1.4 Evaluate the data from Italy

```{r evaluate gdp in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Analyse the different variables for stationarity
analyze_time_series <- function(ts_data, country, variable) {

  # Doing the ADF and PP test, as well as determining diff
  adf_result <- tseries::adf.test(ts_data)
  num_diffs <- ndiffs(ts_data)

  # Print results for the current variable
  cat("Results for", variable, "in", country, ":\n")
  cat("ADF Test:\n")
  cat("  P-Value:", adf_result$p.value, "\n")
  cat("Number of Differences Required for Stationarity:", num_diffs, "\n\n")
  
  # Set up the plot
  par(mfrow = c(1, 2), mar = c(3, 4, 4, 1))

  # ACF and PACF plots
  acf(ts_data, main = paste("ACF of", variable, "in", country))
  pacf(ts_data, main = paste("PACF of", variable, "in", country))
}

# Explore gdp of Italy
analyze_time_series(transformed_italy$gdp, "Italy", 'GDP')

```

In our financial analysis of Italy's GDP, the Autocorrelation Function (ACF) displayed 19 significant lags, as well as the before seans seasonal pattern. Due to the slow decline we might conclude that it is an AR model. This is further supported due to the steap decline in the Partial Autocorrelation Function (PACF) after the fifth lag. Also the Model will probably include a seasonal component. The Augmented Dickey-Fuller (ADF) test has a p- value of 0.26, which is more than 0.05, indicating that the data is non-stationary. The number of differences required for stationarity is 1. 


```{r differenced gdp in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Analyse the differenced variables for stationary
analyze_diff_series <- function(ts_data, country, variable, seasonal = 0) {
  
  # Getting number of differences required for stationarity
  num_diffs <- ndiffs(ts_data)
  
  # Differncing num_diffs times
  for (i in 1:num_diffs) {
    ts_data <- diff(ts_data)
  }
  
  if (seasonal != 0){
    ts_data <- diff(ts_data, lag = seasonal)
  }
  
  # Performing the ADF and PP test
  adf_result <- tseries::adf.test(ts_data)
  num_diffs_new <- ndiffs(ts_data)
  
  # Print results for the current variable
  cat("Results for", variable, "in", country, "with season", seasonal, ":\n")
  cat("ADF Test:\n")
  cat("  P-Value:", adf_result$p.value, "\n")
  cat("Number of Differences Required for Stationarity:", num_diffs_new, "\n\n")

}

# Difference data from Italy
analyze_diff_series(transformed_italy$gdp, "Italy", 'GDP')
analyze_diff_series(transformed_italy$gdp, "Italy", 'GDP', 4)

```

After using the recomended number of differences (1) on Italy's GDP the Augmented Dickey-Fuller (ADF) test suggested marginal evidence against stationarity (p = 0.0675), as well as no further differencing. Since this is not compleatly stationary using the knowledge from the ploted data the seasonal component was added. In the seasonal scenario (season = 4), the ADF test provided strong evidence for stationarity (p = 0.01). This suggests to use one difference in the modeling part, as well as including a seasonal componant with a lag of 4.


# TOOD after this not checked the descrptions, style like above is good

```{r evaluate cpi in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Explore cpi of Italy
analyze_time_series(transformed_italy$cpi, "Italy", 'CPI')

```

In our financial analysis of Italy's Consumer Price Index (CPI), the Autocorrelation Function (ACF) showed a gradual decline, especially at lag 19. This hints at weakening correlations over time. Looking ahead, we expect a noticeable spike in the Partial Autocorrelation Function (PACF), reflecting the declining ACF trend. The confidence intervals of -0.1 and 0.1 add statistical weight to these correlations. This info helps us predict a significant positive spike in the PACF, offering valuable insights for future forecasting and decisions.

TODO Description like in GDP Italy

```{r differenced cpi in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Explore cpi of Italy
analyze_diff_series(transformed_italy$cpi, "Italy", 'CPI')

```

In the analysis of Italy's Consumer Price Index, the Augmented Dickey-Fuller (ADF) test yielded a p-value of 0.0235, and the Phillips-Perron (PP) test produced a p-value of 0.01. These results indicate statistical significance, suggesting that the CPI series is likely stationary without the need for differencing (the number of differences required for stationarity is 0). In conclusion, the CPI data for Italy appears to show stationarity, which is a crucial characteristic for time series analysis, making it more suitable for reliable statistical modeling and forecasting.

TODO Description like in GDP Italy

```{r evaluate stocks in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Explore stocks of Italy
analyze_time_series(transformed_italy$stocks, "Italy", 'Stocks')


```

TODO Description like in GDP Italy

```{r differenced stocks in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Explore stocks of Italy
analyze_diff_series(transformed_italy$stocks, "Italy", 'Stocks')

```

TODO Description like in GDP Italy

## 1.4 Evaluate the data from USA

```{r evaluate gdp in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Explore gdp of USA
analyze_time_series(transformed_usa$gdp, "USA", 'GDP')

```

TODO Description like in GDP Italy

```{r differenced gdp in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Explore gdp of Italy
analyze_diff_series(transformed_usa$gdp, "USA", 'GDP')
analyze_diff_series(diff(transformed_usa$gdp), "USA", 'GDP', 4)

```

Indication of a drift or linear trend that is not diffenced (no clue why), TODO Yannik

TODO Description like in GDP Italy

```{r evaluate cpi in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Explore cpi of USA
analyze_time_series(transformed_usa$cpi, "USA", 'CPI')

```

TODO Description like in GDP Italy

```{r differenced cpi in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Explore cpi of Italy
analyze_diff_series(transformed_usa$cpi, "USA", 'CPI')

```

TODO Description like in GDP Italy

```{r evaluate stocks in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Explore stocks of USA
analyze_time_series(transformed_usa$stocks, "USA", 'Stocks')

```

TODO Description like in GDP Italy

```{r differenced stocks in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Explore stocks of Italy
analyze_diff_series(transformed_usa$stocks, "USA", 'Stocks')

```

TODO Description like in GDP Italy

## 1.5 Volatility clustering

QUESTION? Is that it?

```{r volatility clustering italy gdp, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

plot_volatility_clusters <- function(time_series, country, variable, uper = 0) {
  
  # Calculate the squared returns
  time_series <- diff(time_series)^2
  
  if (uper == 0) {
    uper <- max(time_series) + max(time_series)*0.05
  }

  # Create the line plot
  ggplot(data.frame(time = time(time_series), value = time_series), aes(x=time,y=value))  +
    geom_segment(aes(xend = time, yend = 0)) +
    labs(title = paste("Squared returns of", variable, 'in', country),
         x = "Time",
         y = "Squared returns") + 
    theme_minimal() +
    theme(panel.grid.minor = element_blank(),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5)) +
    scale_y_continuous(limits = c(0, uper))

}

# Plotting the volatility clustering 
plot_volatility_clusters(transformed_italy$stocks, 'Italy', 'GDP')


```


```{r volatility clustering italy cpi, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Plotting the volatility clustering
plot_volatility_clusters(transformed_italy$cpi, 'Italy', 'CPI')

```
QUESTION Heteroskedasticity so should we use GARCH or use log or maybe box cox?

```{r volatility clustering italy stocks, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Plotting the volatility clustering
plot_volatility_clusters(transformed_italy$stocks, 'Italy', 'Stocks')

```

```{r volatility clustering usa gdp, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Plotting the volatility clustering
plot_volatility_clusters(transformed_usa$gdp, 'USA','GDP')


```

```{r volatility clustering usa cpi, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Plotting the volatility clustering
plot_volatility_clusters(transformed_usa$cpi, 'USA', 'CPI', 10)

```



```{r volatility clustering usa stocks, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Plotting the volatility clustering
plot_volatility_clusters(transformed_usa$stocks, 'USA', 'Stocks', 100)

```



<br>

------------------------------------------------------------------------

<br>

# Part 2: Model

------------------------------------------------------------------------

## 2.1 Train test split

QUESTIONHERE2: How should we split the data?

```{r train test split, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Creang train test split function
train_test_split <- function(data_list, train_percentage = 0.8) {
  split_data <- list()
  
  # Loop through the variables in the list
  for (variable in names(data_list)) {
    ts_data <- data_list[[variable]]
    data_length <- length(ts_data)

    # Calculate the train size
    train_size <- round(train_percentage * data_length)

    # Split the data
    train_data <- ts_data[1:train_size]
    test_data <- ts_data[(train_size + 1):data_length]
    
    # Add the data to the list
    split_data[[paste0(variable, "_train")]] <- train_data
    split_data[[paste0(variable, "_test")]] <- test_data
  }

  return(split_data)
}

# Apply the train test split function
split_italy <- train_test_split(transformed_italy)
split_usa <- train_test_split(transformed_usa)

print(is.ts(split_italy$gdp_train)) # TODO IS NOT A TS ANYMORE SHE DID NOT CARE EITHER?
print(split_italy$gdp_train) 


```

Had for some reason huge troubles with it sting in ts so needed to create this

```{r train test split alternative, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Function for train test split
train_test_split <- function(data_list, train_percentage = 0.8) {
  split_data <- list()
  
  # Loop through the variables in the list
  for (variable in names(data_list)) {
    ts_data <- data_list[[variable]]
    data_length <- length(ts_data)

    # Calculate the train size
    train_size <- round(train_percentage * data_length)
    
    # Train split, this way needed, to keep the ts format
    train_data <- ts(ts_data[1:train_size], start = start(ts_data), frequency = frequency(ts_data))

    end_time <- time(train_data)[length(train_data)]
    
    # Get start for testdata, if only split with [:] the ts type is lost as well as the start time
    if (frequency(ts_data) == 12) {
      end_time <- end_time + 1/12
      start_year <- floor(end_time)
      start_month <- (end_time - start_year) * 12 + 1
    } else if (frequency(ts_data) == 4) {
      end_time <- end_time + 1/4
      start_year <- floor(end_time)
      start_month <- (end_time - start_year) * 4 + 1
    } else {
      print('Error: Frequency not supported')
    }

    test_data <- ts(ts_data[(train_size + 1):data_length], start = c(start_year, start_month), frequency = frequency(ts_data))
    
    # Add the data to the list
    split_data[[paste0(variable, "_train")]] <- train_data
    split_data[[paste0(variable, "_test")]] <- test_data
  }

  return(split_data)
}

# Apply the train test split function
split_italy <- train_test_split(transformed_italy)
split_usa <- train_test_split(transformed_usa)

```

```{r extract data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Italy variables
gdp_italy_train <- split_italy$gdp_train
gdp_italy_test <- split_italy$gdp_test

cpi_italy_train <- split_italy$cpi_train
cpi_italy_test <- split_italy$cpi_test

stocks_italy_train <- split_italy$stocks_train
stocks_italy_test <- split_italy$stocks_test

# USA variables
gdp_usa_train <- split_usa$gdp_train
gdp_usa_test <- split_usa$gdp_test

cpi_usa_train <- split_usa$cpi_train
cpi_usa_test <- split_usa$cpi_test

stocks_usa_train <- split_usa$stocks_train
stocks_usa_test <- split_usa$stocks_test

```

## 2.2 Trend and Seasonality with holt???

Only for Italy gdp and cpi, or should we put this after the visual inspection? TODO

```{r stationarity italy cpi, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}


plot_forecast <- function(train_data, test_data, ts_data, country, variable, alpha, model) {
  model_hw <- ets(train_data, model = model, alpha = alpha)
  ntest <- length(test_data)

  pred <- forecast(model_hw, h = ntest, level = 0)

  pred_hw <- c(model_hw$fitted, pred$mean)

  # Extract the last date of the training data
  last_date_train <- tail(time(train_data), 1)

  ggplot(data.frame(time = time(ts_data), value = ts_data, pred_hw = pred_hw)) + 
    geom_line(aes(x = time, y = value, group = 1)) +
    geom_line(aes(x = time, y = pred_hw, group = 1), color = "dodgerblue4", size = 0.75) +
    geom_vline(aes(xintercept = as.numeric(last_date_train)), color = "darkorange2", size = 0.9) +
    xlab("Year") +
    ylab(paste(variable, "in", country)) +
    labs(title = paste("Forecast for", variable, "in", country),
         subtitle = paste("Exponential Smoothing (ETS) with alpha =", alpha),
         caption = "Vertical line represents the last date of the training data") +
    theme_minimal() +
    theme(plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
          plot.subtitle = element_text(size = 10, hjust = 0.5),
          plot.caption = element_text(size = 6))
  
}

plot_forecast(gdp_italy_train, gdp_italy_test, transformed_italy$gdp, country = "Italy", variable = "GDP", alpha = 0.7, 'AAA')

```

```{r stationarity italy gdp, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

plot_forecast(cpi_italy_train, cpi_italy_test, transformed_italy$cpi, country = "Italy", variable = "CPI", alpha = 0.7, 'AAA')


```

## 2.3 ARIMA

TODO should we do combined arch garch? is thata worth it for every or for which?

TO WE NEED THIS REALLY OR DO WE NOT USE IT? QUESTIONHERE3
```{r TODO model automation, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# QUESTION DO WE WANT TO USE THIS? 

create_models <- function(p, d, q, train_data, test_data) {
  results <- data.frame(order = character(),
                         aic = numeric(),
                         bic = numeric(),
                         mpse = numeric())

  for (i in 0:p) {
    for (j in 0:q) {
      order <- c(i, d, j)
      fit <- Arima(train_data, order = order)
      aic <- AIC(fit)
      bic <- BIC(fit)

      pred <- forecast(fit, h = length(test_data))
      mpse <- mean((pred$mean - test_data)^2)

      results <- rbind(results, data.frame(order = paste(order, collapse = "_"),
                                           aic = aic,
                                           bic = bic,
                                           mpse = mpse))
    }
  }

  # Choose parameters for the 3 lowest BIC, 3 lowest AIC, and 3 lowest MPSE
  best_bic <- results[order(results$bic), ][1:3, ]
  best_aic <- results[order(results$aic), ][1:3, ]
  best_mpse <- results[order(results$mpse), ][1:3, ]

  # Extract the selected orders
  selected_orders <- unique(rbind(best_bic, best_aic, best_mpse)$order)

  # Fit models with the selected orders
  selected_models <- lapply(strsplit(selected_orders, "_"), function(x) {
    order <- as.numeric(x)
    fit <- Arima(train_data, order = order)
    return(fit)
  })

  return(selected_models)
}

# Example usage:
selected_models <- create_models(p = 2, d = 1, q = 2, train_data = gdp_italy_train, test_data = gdp_italy_test)

```

### 2.3.1 GDP Italy
```{r modelling italy gdp, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}  
# Function to format ARIMA order
format_arima_order <- function(arma) {
  non_seasonal <- paste(arma[1], arma[6], arma[2])
  seasonal <- paste(arma[3], arma[7], arma[4])
  period <- arma[5]
  return(paste("ARIMA(", non_seasonal, ")", "(", seasonal, ")", "[", period, "]", sep = ""))
}

# Function for model comparison
evaluate_arima_models <- function(models, test_data) {
  
  results <- data.frame(Model = character(), AIC = numeric(), SIC = numeric(), BIC = numeric(), Residuals = character(), MSPE = numeric())
  
  for (model in models) {
    fit <- model
    order <- coef(fit)
    
    # Extract AIC and SIC
    aic <- AIC(fit)
    sic <- AIC(fit, k = log(length(fit$residuals)))
    bic <- BIC(fit)
    
    # Check residuals for white noise
    white_noise <- Box.test(fit$residuals, lag = 20, type = "Ljung-Box")$p.value > 0.05
    residuals_check <- ifelse(white_noise, "White Noise", "Not White Noise")
    
    # Make predictions on test data
    pred <- forecast(fit, h = length(test_data))
    
    # Calculate MSPE
    mspe <- mean(((pred$mean - test_data) / test_data)^2, na.rm = TRUE)*100


    # Append results to the data frame
    name <- format_arima_order(model$arma)
    result_row <- data.frame(Model = name, AIC = aic, SIC = sic, BIC = bic, Residuals = residuals_check, MSPE = mspe)
    results <- rbind(results, result_row)
  }

  return(results)
}


# TODO Use the seasonal knowledge tocreate one ourselfs

# Choose the maximum lags
model_max <- forecast::Arima(gdp_italy_train, order = c(9, 1, 16), method = 'ML')

# Choose the best model based on AIC
model_aic <- auto.arima(gdp_italy_train, ic = 'aic')

# Choose the best model based on BIC
model_bic <- auto.arima(gdp_italy_train, ic = 'bic')


# Create a list of models
model_list <- list(model_max, model_aic, model_bic)

# Evaluate the models
evaluate_arima_models(model_list, gdp_italy_test)

```  

```{r plot model italy gdp, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}  

# Plot the forecast
plot_forecast_arima <- function(train_data, test_data, ts_data, country, variable, model) {
  ntest <- length(test_data)
  pred <- forecast(model, h = ntest, level = 0)
  pred_arima <- c(model$fitted, pred$mean)
  
  # Extract the last date of the training data
  last_date_train <- tail(time(train_data), 1)
  
  name <- format_arima_order(model$arma)
  
  ggplot(data.frame(time = time(ts_data), value = ts_data, pred_arima = pred_arima)) + 
    geom_line(aes(x = time, y = value, group = 1)) +
    geom_line(aes(x = time, y = pred_arima, group = 1), color = 'darkgreen', size=0.75) +
    geom_vline(aes(xintercept = as.numeric(last_date_train)), color = "darkblue", size = 0.9) +
    xlab("Year") +
    ylab(paste(variable, "in", country)) +
    labs(title = paste("Forecast for", variable, "in", country),
         subtitle = paste("The best model is", name),
         caption = "Vertical line represents the last date of the training data") +
    theme_minimal() +
    theme(plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
          plot.subtitle = element_text(size = 10, hjust = 0.5),
          plot.caption = element_text(size = 6))
}

plot_residuals_arima <- function(model) { # QUESTION This is a good function to evaluate the residuals? and why is it white noise but ther are still iolations in the model_bic?
  residuals <- model$residuals
  acf(residuals, lag.max = 20, main = "ACF of Residuals")
}


# Plot the best model
plot_forecast_arima(gdp_italy_train, gdp_italy_test, transformed_italy$gdp, "Italy", "GDP", model_bic)
plot_residuals_arima(model_bic)
# QUESTION why is it white nois adn there is still a violation?

```


### 2.3.1 CPI Italy

```{r modelling italy cpi, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# TODO maybe look at seasonal self composed once?

# Creating different models
model_med <- forecast::Arima(cpi_italy_train, order = c(7, 1, 0)) 
model_aic <- auto.arima(cpi_italy_train, ic = 'aic')
model_bic <- auto.arima(cpi_italy_train, ic = 'bic')

# Create a list of models
model_list <- list(model_max, model_med, model_aic, model_bic)

# Evaluate the models
evaluate_arima_models(model_list, cpi_italy_test)

```  

```{r plot model italy cpi, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Trying other models (GARCH?)
```  



TODO HERE WE NEED TO LOOK INTO THE BEST MODELS BECAUSE IT IS NOT ONLY WHITE NOISE with 
Plot residuals
plot(model_max$residuals)
acf(model_aic$residuals)
pacf(model_aic$residuals)

```{r plot modelling italy cpi, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot the forecast
plot_forecast_arima(cpi_italy_train, cpi_italy_test, transformed_italy$cpi, "Italy", "CPI", model_aic)

```  

### 2.3.1 Stocks Italy

```{r modelling italy stocks, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# TODO maybe directly GARCH?

# Creating different models
model_min <- forecast::Arima(stocks_italy_train, order = c(2, 1, 0)) 
model_aic <- auto.arima(stocks_italy_train, ic = 'aic')
model_bic <- auto.arima(stocks_italy_train, ic = 'bic')

# Create a list of models
model_list <- list(model_min, model_aic, model_bic)

# Evaluate the models
evaluate_arima_models(model_list, stocks_italy_test)

```  

TODO this looks super bad doesnt it? So GARCH needed but WHY? not visible in the plots was it?

### 2.3.1 GDP USA

```{r modelling usa gdp, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Creating different models
model_min <- forecast::Arima(gdp_usa_train, order = c(1, 1, 0))
model_aic <- auto.arima(gdp_usa_train, ic = 'aic')
model_bic <- auto.arima(gdp_usa_train, ic = 'bic')

# Create a list of models
model_list <- list(model_min, model_aic, model_bic)

# Evaluate the models
evaluate_arima_models(model_list, gdp_usa_test)

``` 

### 2.3.1 CPI USA

```{r modelling usa cpi, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Creating different models
model_min <- forecast::Arima(cpi_usa_train, order = c(1, 1, 0))
model_aic <- auto.arima(cpi_usa_train, ic = 'aic')
model_bic <- auto.arima(cpi_usa_train, ic = 'bic')

# Create a list of models
model_list <- list(model_min, model_aic, model_bic)

# Evaluate the models
evaluate_arima_models(model_list, cpi_usa_test)

# TODO why, isnt the PCA 1 and that is it?

``` 

### 2.3.1 Stocks USA

```{r modelling usa stocks, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Creating different models
model_min <- forecast::Arima(stocks_usa_train, order = c(1, 1, 0))
model_max <- forecast::Arima(stocks_usa_train, order = c(6, 1, 0))
model_aic <- auto.arima(stocks_usa_train, ic = 'aic')
model_bic <- auto.arima(stocks_usa_train, ic = 'bic')

# Create a list of models
model_list <- list(model_min, model_max, model_aic, model_bic)

# Evaluate the models
evaluate_arima_models(model_list, stocks_usa_test)

``` 
