---
title: "Advanced Data Analytics"
author:
  - Loukas Gonzalez
  - Simon Alvarez Bliffeld
  - Simon Gonzalez
  - Yannik Biebert
date: "2023-11-28"
output:      
  rmarkdown::html_document:
    theme: cosmo
    highlight: tango
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
---

```{r setup, include=FALSE}
# Set up: do not change!!!
knitr::opts_chunk$set(echo = TRUE)
```

<br>

------------------------------------------------------------------------

<br>

# Part 1: Prepare and inspect the data

Steps for us:

- Visual inspection of data DONE

- Stationarity testing DONE

- Inspect autocorrelation with acf and pacf DONE

- Inspect volatility clustering (if so, maybe GARCH needed?) 

- Inspect trend and seasonality with zoom in or decompose and if it has use holt-winters, DO WE NEED TBATS or STLM? Because I dont think whe have two cicles in each other

- Maybe new autocorrelation with acf and pacf for the estimated seansonality and trend (if we use the holt winter model to mode arima)

- Look for ARIMA models (on data and on estimated ets()/msts()

- Identify the seasonal model by analyzing the seasonal coefficients of the ACF and PACF

- Identify the regular component by exploring the ACF and PACF of the residuals of the seasonal model.

- Compare models (including ets models)- AIC or Accuracy Metrics

- Choose best model and visualize, look at residuals, maybe GARCH (or should be aready dicide this before using ARIMA Model and use it directly?)

- Use maybe Garch, and compare those models

- Summary of findings

- Forcast and plot final model

- conclusion of forecast

------------------------------------------------------------------------

## 1.1 Importing and inspecting the data set

```{r Import data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# import packages
library(pacman)
p_load(
  readxl, tseries, tidyverse, dplyr, tidyr, forecast, fpp2, stats, quantmod, zoo, lubridate, ggplot2, cowplot
)

# import data set xlxs
data <- read_excel("Datasets.xlsx")

# removing not used columns Housing and Unemployment
data <- data[, -c(7, 9)]


# Display the data set
print(data)

```

TODO Description

## 1.2 Prepare the data

```{r transform data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Split the data into two dataframes based on the country
data_italy <- subset(data, Country %in% c("Italy"))
data_usa <- subset(data, Country %in% c("USA"))

# Function for data transformation
transform_data <- function(df) {
  # Extract columns from the dataframe
  gdp <- df$'GDP, billion currency units'
  cpi <- df$'Consumer Price Index (CPI)'
  stocks <- df$'Stock market index'

  # Drop missing values (becuase only every quarter)
  gdp <- na.omit(gdp)

  # Drop missing values (becuase last in the month)
  cpi <- na.omit(cpi)
  stocks <- na.omit(stocks)

  # Change to time series
  gdp <- ts(gdp, start = c(2000, 1), frequency = 4)
  cpi <- ts(cpi, start = c(2000, 1), frequency = 12)
  stocks <- ts(stocks, start = c(2000, 1), frequency = 12)
  
  # Return the time series as separate variables
  return(list(gdp = gdp, cpi = cpi, stocks = stocks))
}

# Apply the transformation function
transformed_italy  <- transform_data(data_italy)
transformed_usa <- transform_data(data_usa)

```

TODO Description

## 1.3 Visual inspection of the data, 1.6 Trend and Seasonality here?

```{r plot gdp in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Create a function for visualization.
plot_and_zoom <- function(data, num_years) {
  # Extract the specified number of years from the time series
  end_year <- start(data)[1] + num_years - 1
  window_data <- window(data, start = start(data), end = c(end_year, 12))
  
  # Create the main plot
  main_plot <- ggplot(data, aes(x = time(data), y = as.numeric(data))) +
    geom_line() +
    theme_minimal() +
    labs(title = "GDP in Italy", x = "Year", y = "GDP") +
    theme(panel.grid.minor = element_blank(),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
  
  # Create a zoomed-in version with the specified number of years
  zoomed_in_plot <- ggplot(window_data, aes(x = time(window_data), y = as.numeric(window_data))) +
    geom_line() +
    labs(title = "Zommed in GDP in Italy", x = "Year", y = "GDP") +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(),
      plot.title = element_text(size = 12, face = "bold", hjust = 0.5))
  
  plot_grid(main_plot, zoomed_in_plot, ncol = 1)
  
}

# Call the function with your time series data
plot_and_zoom(transformed_italy$gdp, 4)

```

TODO Description

```{r plot cpi in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot the Italy cpi
plot(transformed_italy$cpi, main = "CPI in Italy"
     , xlab = "Year", ylab = "CPI")

```

TODO Description

```{r plot stocks in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot the Italy data
plot(transformed_italy$stocks, main = "Stocks in Italy"
     , xlab = "Year", ylab = "Stocks")

```

TODO Description

```{r plot gdp in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot the USA gdp
plot(transformed_usa$gdp, main = "GDP in the USA"
     , xlab = "Year", ylab = "GDP")

```

TODO Description

```{r plot cpi in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot the USA cpi
plot(transformed_usa$cpi, main = "CPI in the USA"
     , xlab = "Year", ylab = "CPI")

```

TODO Description

```{r plot stocks in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Plot the USA stocks
plot(transformed_usa$stocks, main = "Stocks in the USA"
     , xlab = "Year", ylab = "Stocks")

```

TODO Describe

## 1.4 Evaluate the data from Italy

```{r evaluate gdp in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Analyse the different variables for stationarity
analyze_time_series <- function(ts_data, country, variable) {

  # Doing the ADF and PP test, as well as determining diff
  adf_result <- adf.test(ts_data)
  pp_result <- pp.test(ts_data)
  num_diffs <- ndiffs(ts_data)

  # Print results for the current variable
  cat("Results for", variable, "in", country, ":\n")
  cat("ADF Test:\n")
  cat("  P-Value:", adf_result$p.value, "\n")
  cat("PP Test:\n")
  cat("  P-Value:", pp_result$p.value, "\n")
  cat("Number of Differences Required for Stationarity:", num_diffs, "\n\n")
  
  # Set up the plot
  par(mfrow = c(1, 2), mar = c(3, 4, 4, 1))

  # ACF and PACF plots
  acf(ts_data, main = paste("ACF of", variable, "in", country))
  pacf(ts_data, main = paste("PACF of", variable, "in", country))
}

# Explore gdp of Italy
analyze_time_series(transformed_italy$gdp, "Italy", 'GDP')

```

TODO Description

```{r differenced gdp in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Analyse the differenced variables for stationary
analyze_diff_series <- function(ts_data, country, variable) {
  
  # Getting number of differences required for stationarity
  num_diffs <- ndiffs(ts_data)
  
  # Differncing num_diffs times
  for (i in 1:num_diffs) {
    ts_data <- diff(ts_data)
  }
  
  # Performing the ADF and PP test
  adf_result <- adf.test(ts_data)
  pp_result <- pp.test(ts_data)
  num_diffs_new <- ndiffs(ts_data)
  
  # Print results for the current variable
  cat("Results for", variable, "in", country, ":\n")
  cat("ADF Test:\n")
  cat("  P-Value:", adf_result$p.value, "\n")
  cat("PP Test:\n")
  cat("  P-Value:", pp_result$p.value, "\n")
  cat("Number of Differences Required for Stationarity:", num_diffs_new, "\n\n")

}

# Difference data from Italy
analyze_diff_series(transformed_italy$gdp, "Italy", 'GDP')

```

TODO Description

```{r evaluate cpi in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Explore cpi of Italy
analyze_time_series(transformed_italy$cpi, "Italy", 'CPI')

```

TODO Description

```{r differenced cpi in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Explore cpi of Italy
analyze_diff_series(transformed_italy$cpi, "Italy", 'CPI')

```

TODO Description

```{r evaluate stocks in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Explore stocks of Italy
analyze_time_series(transformed_italy$stocks, "Italy", 'Stocks')


```

TODO Description

```{r differenced stocks in italy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Explore stocks of Italy
analyze_diff_series(transformed_italy$stocks, "Italy", 'Stocks')

```

TODO Description


## 1.4 Evaluate the data from USA

```{r evaluate gdp in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Explore gdp of USA
analyze_time_series(transformed_usa$gdp, "USA", 'GDP')

```

TODO Description

```{r differenced gdp in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Explore gdp of Italy
analyze_diff_series(transformed_usa$gdp, "USA", 'GDP')

```

TODO Description

```{r evaluate cpi in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Explore cpi of USA
analyze_time_series(transformed_usa$cpi, "USA", 'CPI')

```

TODO Description

```{r differenced cpi in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Explore cpi of Italy
analyze_diff_series(transformed_usa$cpi, "USA", 'CPI')

```

TODO Description

```{r evaluate stocks in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Explore stocks of USA
analyze_time_series(transformed_usa$stocks, "USA", 'Stocks')

```

TODO Description

```{r differenced stocks in usa, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Explore stocks of Italy
analyze_diff_series(transformed_usa$stocks, "USA", 'Stocks')

```

TODO Description

## 1.5 Volatility clustering

TODO? Is that it?

```{r volatility clustering italy gdp, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

plot_volatility_clusters <- function(time_series, variable, country,width = 4) {
  
  # Calculate the rolling standard deviation
  rolling_sd <- zoo::rollapply(time_series, FUN = sd, align = "center", width = width)

  # Assuming rolling_sd is already calculated
  rolling_sd_data <- data.frame(Time = time(rolling_sd), RollingSD = coredata(rolling_sd))
  
  # Create the line plot
  ggplot(data = rolling_sd_data, aes(x = Time, y = RollingSD)) +
    geom_line() +
    labs(title = paste("Rolling Standard Deviation of", variable, 'in', country),
         x = "Time",
         y = "Rolling Standard Deviation")
}

# Plotting the volatility clustering 
plot_volatility_clusters(transformed_italy$stocks, 'Italy', 'GDP', 8) # TODO which width?


```

TODO Descripiton

```{r volatility clustering italy cpi, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Plotting the volatility clustering
plot_volatility_clusters(transformed_italy$cpi, 'Italy', 'CPI', 8)

```

TODO Descripiton

```{r volatility clustering italy stocks, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Plotting the volatility clustering
plot_volatility_clusters(transformed_italy$stocks, 'Italy', 'Stocks', 8)

```

TODO Descripiton

```{r volatility clustering usa gdp, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Plotting the volatility clustering
plot_volatility_clusters(transformed_usa$gdp, 'USA','GDP', 8)


```

TODO Descripiton

```{r volatility clustering usa cpi, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Plotting the volatility clustering
plot_volatility_clusters(transformed_usa$cpi, 'USA', 'CPI', 8)

```

TODO Descripiton

```{r volatility clustering usa stocks, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}

# Plotting the volatility clustering
plot_volatility_clusters(transformed_usa$stocks, 'USA', 'Stocks', 8)

```

TODO Descripiton

## 1.6 Trend and Seasonality

Only for Italy gdp?, or should we put this after the visual inspection? TODO

```{r stationarity italy gdp, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=8, fig.height=3}
# Zoom into the stock and buil the ets()


```



<br>

------------------------------------------------------------------------

<br>

# Part 2: Model


------------------------------------------------------------------------


## 2.1 Train test split

```{r train test split, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Creang train test split function
train_test_split <- function(data_list, train_percentage = 0.8) {
  split_data <- list()
  
  # Loop through the variables in the list
  for (variable in names(data_list)) {
    ts_data <- data_list[[variable]]
    data_length <- length(ts_data)

    # Calculate the train size
    train_size <- round(train_percentage * data_length)

    # Split the data
    train_data <- ts_data[1:train_size]
    test_data <- ts_data[(train_size + 1):data_length]
    
    # Add the data to the list
    split_data[[paste0(variable, "_train")]] <- train_data
    split_data[[paste0(variable, "_test")]] <- test_data
  }

  return(split_data)
}

# Apply the train test split function
split_italy <- train_test_split(transformed_italy)
split_usa <- train_test_split(transformed_usa)

print(is.ts(split_italy$gdp_train)) # TODO IS NOT A TS ANYMORE SHE DID NOT CARE EITHER?
print(split_italy$gdp_train) 


```

Had for some reason huge troubles with it sting in ts so needed to create this

```{r train test split alternative, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# TODO Do we want to use this or the one above?

# Function for train test split
train_test_split <- function(data_list, train_percentage = 0.8) {
  split_data <- list()
  
  # Loop through the variables in the list
  for (variable in names(data_list)) {
    ts_data <- data_list[[variable]]
    data_length <- length(ts_data)

    # Calculate the train size
    train_size <- round(train_percentage * data_length)
    
    # Train split, this way needed, to keep the ts format
    train_data <- ts(ts_data[1:train_size], start = start(ts_data), frequency = frequency(ts_data))

    end_time <- time(train_data)[length(train_data)]
    
    # Get start for testdata, if only split with [:] the ts type is lost as well as the start time
    if (frequency(ts_data) == 12) {
      end_time <- end_time + 1/12
      start_year <- floor(end_time)
      start_month <- (end_time - start_year) * 12 + 1
    } else if (frequency(ts_data) == 4) {
      end_time <- end_time + 1/4
      start_year <- floor(end_time)
      start_month <- (end_time - start_year) * 4 + 1
    } else {
      print('Error: Frequency not supported')
    }

    test_data <- ts(ts_data[(train_size + 1):data_length], start = c(start_year, start_month), frequency = frequency(ts_data))
    
    # Add the data to the list
    split_data[[paste0(variable, "_train")]] <- train_data
    split_data[[paste0(variable, "_test")]] <- test_data
  }

  return(split_data)
}

# Apply the train test split function
split_italy <- train_test_split(transformed_italy)
split_usa <- train_test_split(transformed_usa)

```


```{r extract data, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}

# Italy variables
gdp_italy_train <- split_italy$gdp_train
gdp_italy_test <- split_italy$gdp_test

cpi_italy_train <- split_italy$cpi_train
cpi_italy_test <- split_italy$cpi_test

stocks_italy_train <- split_italy$stocks_train
stocks_italy_test <- split_italy$stocks_test

# USA variables
gdp_usa_train <- split_usa$gdp_train
gdp_usa_test <- split_usa$gdp_test

cpi_usa_train <- split_usa$cpi_train
cpi_usa_test <- split_usa$cpi_test

stocks_usa_train <- split_usa$stocks_train
stocks_usa_test <- split_usa$stocks_test

```

## 2.2 Italy

### 2.2.1 GDP

```{r modelling italy gdp, echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE}
# Function for model creation
ntest <- length(gdp_italy_test)

# TODO in text : ACF slow decay, PACF sharp, probably AR 
# 16 MA
# Diff 1
# 9 AR

model_max <- Arima(gdp_italy_train, order = c(9, 1, 16), method = "ML")
model_aic <- auto.arima(gdp_italy_train, ic = 'aic')
model_bic <- auto.arima(gdp_italy_train, ic = 'bic')

predict_max <- forecast::forecast(model_max, h = ntest, level = 0)
predict_aic <- forecast::forecast(model_aic, h = ntest, level = 0)
predict_bic <- forecast::forecast(model_bic, h = ntest, level = 0)

# Get the AIC 

################## HAVE TO WORK WITH cat()

model_max$aic
model_aic$aic
model_bic$aic

print(' ')

accuracy(predict_max, gdp_italy_test)
accuracy(predict_aic, gdp_italy_test)
accuracy(predict_bic, gdp_italy_test)

# Two best:
model_aic # WHAT IS SMA1?
model_bic


# Box.test(auto_model_adjusted_bic$residuals) # p-value = 0.01 --> Residuals NOT white noise
                                            # Big p is only white noise


```


```{r}

### plot the forecasts

# elp$pred <- c(model$fitted, pred$mean)

# ggplot(elp)+ 
#   geom_line(aes(mdy(DATE), Value, group = 1))+
#   geom_line(aes(mdy(DATE), pred, group = 1), color = "dodgerblue4")+
#   geom_vline(aes(xintercept=mdy("02-01-2016")), color = "darkorange2", size = 0.9)+
#   xlab("Date")+
#   ylab("Electric production")

### zoom into the test set

# ggplot(elp[374:397,])+
#   geom_line(aes(mdy(DATE), Value, group = 1))+
#   geom_line(aes(mdy(DATE), pred, group = 1), color = "dodgerblue4")+
#   xlab("Date")+
#   ylab("Electric production")


```




